Estimating Life-Stage Specific Habitat Capacity for Anadromous Salmonids using Quantile Random Forest Models

### 2020-01-30 analysis/paper/paper.Rmd
test commit from Sarah's machine. Added comment at start of discussion.
one more test, the first push required credentials
Tried using pander instead of kable to make tables, had to figure out how to get the cross-referenced labels into the pander function (doesn't do it by default, like kable)

---
title: "Estimating Life-Stage Specific Habitat Capacity for Anadromous Salmonids using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah Hoffman:
      email: Sarah.Hoffman@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
abstract: |
  Needs more...

@@

```{r packages}
library(survey)
library(tidyverse)
# library(QRFcapacity) # for the moment, until we know exactly what data to copy over to this repo
library(QRFpaper)
library(janitor)
library(knitr)
library(magrittr)
library(sf)

@@

@@

```{r fh-samp-table}
qrf_data %>%
  # filter(Year <= 2014) %>%
  tabyl(Watershed) %>%
  adorn_totals() %>%
  adorn_pct_formatting() %>%
  # select(-percent) %>%
  rename(`n Sites` = n,
         Percent = percent) %>%
  kable(caption = 'The number of unique sites, by watershed, with paired fish-habitat data used to populate the spring/summer Chinook salmon parr capacity QRF model')
```

```{r select-mets-table}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  mutate(Rank = 1:n()) %>%
  select(-imp) %>%
  left_join(hab_dict) %>%
  filter(MetricGroupName != 'Tier 1 Summary') %>%
  select(Rank,
         Metric = Name,
         `Metric Category` = MetricCategory,
         Description = DescriptiveText) %>%
  kable(caption = "Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.",
        format.args = list(width = c(2, 10, 10, 50)))
```

@@

```{r gaa-covars}
tibble(Covariate = gaa_covars) %>%
  mutate(Scale = if_else(Covariate %in% c('NatPrin1', 
                                          'DistPrin1'),
                         'Watershed (HUC12)',
                         if_else(Covariate %in% c('WIDE_BF',
                                                  'Channel_Type',
                                                  'Elev_M'),
                                 'Site (300 m)',
                                 'Reach (2 km)'))) %>%
  kable(caption = "Globally available attribute (GAA) habitat covariates used to extrapolate quantile regression forest (QRF) model predictions of spring/summer Chinook parr capacity to a larger scale (e.g., watershed, population).")
```

```{r extrap-r2-table}
model_svy_df %>%
  ungroup() %>%
  mutate(n_full = map_dbl(data,
                          .f = nrow),
         k_full = map_dbl(mod_full,
                          .f = function(x) {
                            names(x$coefficients)[!grepl('Intercept', names(x$coefficients))] %>%
                              length
                          }),
         r_full = map_dbl(mod_full,
                          .f = function(x) {
                            cor(x$y, x$fitted.values)
                          }),
         r2_full = r_full^2,
         r2_adj_full = 1 - ( ( (1 - r2_full) * (n_full - 1) ) / (n_full - k_full - 1) )) %>%
  mutate(n_no_champ = map_dbl(data,
                              .f = nrow),
         k_no_champ = map_dbl(mod_no_champ,
                              .f = function(x) {
                                names(x$coefficients)[!grepl('Intercept', names(x$coefficients))] %>%
                                  length
                              }),
         r_no_champ = map_dbl(mod_no_champ,
                              .f = function(x) {
                                cor(x$y, x$fitted.values)
                              }),
         r2_no_champ = r_no_champ^2,
         r2_adj_no_champ = 1 - ( ( (1 - r2_no_champ) * (n_no_champ - 1) ) / (n_no_champ - k_no_champ - 1) )) %>%
  select(response, starts_with("r2")) %>%
  gather(type, r2, -response) %>%
  mutate(model = if_else(grepl('no_champ', type),
                         'non-CHaMP',
                         'CHaMP'),
         type = if_else(grepl('adj', type),
                        'Adjusted R2',
                        'R2')) %>%
  spread(type, r2) %>%
  mutate(response = recode(response,
                           'chnk_per_m' = 'fish/m',
                           'chnk_per_m2' = 'fish/m$^2$')) %>%
  select(Model = model,
         Response = response,
         `$r^2$` = R2, 
         `Adjusted $r^2$` = `Adjusted R2`) %>%
  kable(caption = "Summary of extrapolation model fits, split by each of the responses and whether the extrapolation model used CHaMP watershed as a covariate or not.",
        escape = F,
        digits = c(NA, NA, 3, 3))

@@

```{r sr-table}
spawn_rec_params %>%
  select(Population, data) %>%
  unnest(data) %>%
  group_by(Population) %>%
  summarise(`n Yrs` = n_distinct(BroodYear),
            Spawner_type = unique(Spawner_type)) %>%
  select(Population, `n Yrs`, Spawner_type) %>%
  mutate(Spawner_type = recode(Spawner_type,
                               'total spawners' = 'Spawners',
                               'redd counts' = 'Redds')) %>%
  rename(`Adult Data` = Spawner_type) %>%
  mutate(`Parr Data` = if_else(grepl('Chiwawa', Population),
                               'Parr Surveys',
                               'RST')) %>%
  full_join(spawn_rec_params %>%
              mutate(prnt_val = paste0(prettyNum(round(est), big.mark = ','), ' (', round(cv, 3), ')')) %>%
              select(Population, form, prnt_val) %>%
              mutate(form = recode(form,
                                   'BevertonHolt' = 'Beverton Holt',
                                   'Hockey' = 'Hockey Stick')) %>%
              mutate(form = factor(form,
                                   levels = c('Beverton Holt',
                                              'Ricker',
                                              'Hockey Stick',
                                              'QRF'))) %>%
              spread(form, prnt_val)) %>%
  filter(Population %in% keep_wtsds) %>%
  kable(caption = "Estimates of parr capacity from both spawner-recruit data (Beverton-Holt, Ricker, hockey stick) and from extrapolated estimates of parr capacity from the quantile regression forest (QRF) model. Numbers in parentheses are coefficients of variation.")

>>>>

---
title: "Estimating Life-Stage Specific Habitat Capacity for Anadromous Salmonids using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah Hoffman:
      email: Sarah.Hoffman@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::pdf_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
abstract: |
  Needs more...

@@

```{r packages}
library(survey)
library(tidyverse)
# library(QRFcapacity) # for the moment, until we know exactly what data to copy over to this repo
library(QRFpaper)
library(janitor)
library(knitr)
library(pander)
library(magrittr)
library(sf)

@@

panderOptions("big.mark", ',')

options(knitr.kable.NA = '-')
options(knitr.table.format = "pandoc")

@@

```{r }
qrf_data %>%
  tabyl(Watershed) %>%
  adorn_totals() %>%
  adorn_pct_formatting() %>%
  rename(`n Sites` = n,
         Percent = percent) %>%
  pander(caption = "(\\#tab:fh-samp-table)The number of unique sites, by watershed, with paired fish-habitat data used to populate the spring/summer Chinook salmon parr capacity QRF model.")
  # kable(caption = 'The number of unique sites, by watershed, with paired fish-habitat data used to populate the spring/summer Chinook salmon parr capacity QRF model',
  #       booktabs = T)

```{r select-mets-table}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  mutate(Rank = 1:n()) %>%
  select(-imp) %>%
  left_join(hab_dict) %>%
  filter(MetricGroupName != 'Tier 1 Summary') %>%
  select(Rank,
         Metric = Name,
         `Metric Category` = MetricCategory,
         Description = DescriptiveText) %>%
  pander(caption = "(\\#tab:select-mets-table) Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.")
  # kable(caption = "Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.",
  #       booktabs = T,
  #       format.args = list(width = c(2, 10, 10, 50)))

@@

```{r gaa-covars}
tibble(Covariate = gaa_covars) %>%
  mutate(Scale = if_else(Covariate %in% c('NatPrin1', 
                                          'DistPrin1'),
                         'Watershed (HUC12)',
                         if_else(Covariate %in% c('WIDE_BF',
                                                  'Channel_Type',
                                                  'Elev_M'),
                                 'Site (300 m)',
                                 'Reach (2 km)'))) %>%
  pander(caption = "(\\#tab:gaa-covars)Globally available attribute (GAA) habitat covariates used to extrapolate quantile regression forest (QRF) model predictions of spring/summer Chinook parr capacity to a larger scale (e.g., watershed, population).")
  # kable(caption = "Globally available attribute (GAA) habitat covariates used to extrapolate quantile regression forest (QRF) model predictions of spring/summer Chinook parr capacity to a larger scale (e.g., watershed, population).",
  #       booktabs = T)
```

```{r extrap-r2-table}
model_svy_df %>%
  ungroup() %>%
  mutate(n_full = map_dbl(data,
                          .f = nrow),
         k_full = map_dbl(mod_full,
                          .f = function(x) {
                            names(x$coefficients)[!grepl('Intercept', names(x$coefficients))] %>%
                              length
                          }),
         r_full = map_dbl(mod_full,
                          .f = function(x) {
                            cor(x$y, x$fitted.values)
                          }),
         r2_full = r_full^2,
         r2_adj_full = 1 - ( ( (1 - r2_full) * (n_full - 1) ) / (n_full - k_full - 1) )) %>%
  mutate(n_no_champ = map_dbl(data,
                              .f = nrow),
         k_no_champ = map_dbl(mod_no_champ,
                              .f = function(x) {
                                names(x$coefficients)[!grepl('Intercept', names(x$coefficients))] %>%
                                  length
                              }),
         r_no_champ = map_dbl(mod_no_champ,
                              .f = function(x) {
                                cor(x$y, x$fitted.values)
                              }),
         r2_no_champ = r_no_champ^2,
         r2_adj_no_champ = 1 - ( ( (1 - r2_no_champ) * (n_no_champ - 1) ) / (n_no_champ - k_no_champ - 1) )) %>%
  select(response, starts_with("r2")) %>%
  gather(type, r2, -response) %>%
  mutate(model = if_else(grepl('no_champ', type),
                         'non-CHaMP',
                         'CHaMP'),
         type = if_else(grepl('adj', type),
                        'Adjusted R2',
                        'R2')) %>%
  spread(type, r2) %>%
  mutate(response = recode(response,
                           'chnk_per_m' = 'fish/m',
                           'chnk_per_m2' = 'fish/m$^2$')) %>%
  select(Model = model,
         Response = response,
         `$r^2$` = R2, 
         `Adjusted $r^2$` = `Adjusted R2`) %>%
  pander(caption = "(\\#tab:extrap-r2-table)Summary of extrapolation model fits, split by each of the responses and whether the extrapolation model used CHaMP watershed as a covariate or not.")
  # kable(caption = "Summary of extrapolation model fits, split by each of the responses and whether the extrapolation model used CHaMP watershed as a covariate or not.",
  #       escape = F,
  #       booktabs = T,
  #       digits = c(NA, NA, 3, 3))

@@

```{r sr-table}
spawn_rec_params %>%
  select(Population, data) %>%
  unnest(data) %>%
  group_by(Population) %>%
  summarise(`n Yrs` = n_distinct(BroodYear),
            Spawner_type = unique(Spawner_type)) %>%
  select(Population, `n Yrs`, Spawner_type) %>%
  mutate(Spawner_type = recode(Spawner_type,
                               'total spawners' = 'Spawners',
                               'redd counts' = 'Redds')) %>%
  rename(`Adult Data` = Spawner_type) %>%
  mutate(`Parr Data` = if_else(grepl('Chiwawa', Population),
                               'Parr Surveys',
                               'RST')) %>%
  full_join(spawn_rec_params %>%
              mutate(prnt_val = paste0(prettyNum(round(est), big.mark = ','), ' (', round(cv, 3), ')')) %>%
              select(Population, form, prnt_val) %>%
              mutate(form = recode(form,
                                   'BevertonHolt' = 'Beverton Holt',
                                   'Hockey' = 'Hockey Stick')) %>%
              mutate(form = factor(form,
                                   levels = c('Beverton Holt',
                                              'Ricker',
                                              'Hockey Stick',
                                              'QRF'))) %>%
              spread(form, prnt_val)) %>%
  filter(Population %in% keep_wtsds) %>%
  pander(caption = "(\\#tab:sr-table)Estimates of parr capacity from both spawner-recruit data (Beverton-Holt, Ricker, hockey stick) and from extrapolated estimates of parr capacity from the quantile regression forest (QRF) model. Numbers in parentheses are coefficients of variation.")
  # kable(caption = "Estimates of parr capacity from both spawner-recruit data (Beverton-Holt, Ricker, hockey stick) and from extrapolated estimates of parr capacity from the quantile regression forest (QRF) model. Numbers in parentheses are coefficients of variation.")
